---
title: "Assingment"
author: "Raghu Annavarapu"
date: "02/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Question:1
Describe missing data, provide summary of missing data, similar to the analysis in the Chapter 2 (table 3): Count of missing data/percent per variable, type of missing data (NA, null), total percent of missingness per dataset.

Ans:

Missing Data  is information not available for a subject (or case) about whom other information is available.  Typically occurs when respondent fails to answer one or more questions in a Systematic or Random survey.

Researcher’s Concern is to identify the patterns and relationships underlying the missing data in order to maintain as close as possible to the original distribution of values when any remedy is applied.   

The main impact is it reduces sample size available for analysis and it can distort results.

```{r}

# read csv files
my_data <- read.csv("C:\\Users\\Dell\\Desktop\\Assignment\\weo-data.csv",header = TRUE)
#install naniar: Data Structures, Summaries, and Visualisations for Missing Data
library(naniar)
miss_var_summary(my_data, order = FALSE, add_cumsum = FALSE)
```

Considered world economic indicators data for missing data analysis. From the main dataset, considered data from year 2017 to 2022 for 10 selected countries.Also, removed the variales which are not necessary.

Observations: From years data 5% to 7% data is missing. Overall, 4% data is missing from the dataset. Missing data by Country is also provided for better analysis.
```{r}
# Missing Values Summary - by Country
# works with group_by from dplyr
library(dplyr)
my_data %>%
 group_by(Country) %>%
 miss_var_summary()
```

##  Question:2
Plot visualization of missing data pattern 

Ans:
```{r}
vis_miss(my_data)
vis_miss(my_data, cluster = TRUE)
```

##  Question:3
Describe if you have observed any patterns  

Ans: Overall there are 441 rows of data. In that majorly year data is missing in 23 to 32 rows.

```{r}
library(ggplot2)
gg_miss_var(my_data)
gg_miss_case(my_data)
gg_miss_fct(x = my_data, fct = Country)


ggplot(data = my_data,
       aes(x = WEO.Subject.Code,
           y = Country)) + 
  geom_miss_point() +
  theme(legend.position = "bottom")

```

##  Question:4
Run statistical analysis to determine if your data is MCAR or MAR
```{r}
library(BaylorEdPsych)
library(mvnmle)
LittleMCAR(my_data)
```
##  Question:5
Explain what type of imputation will be performed: list-wise/pair-wise deletions, mean imputation,  regression imputation etc  
```{r}
head(impute_below_if(my_data, is.numeric))
bind_shadow(my_data)
bind_shadow(my_data) %>%
impute_below_all()
```
